---
title: "R Notebook"
author: "Andjanie Somaidien"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---
Data
```{r}
setwd("~/Documents/Tilly Uni/Complex Systems/R")
Plant_1_Generation_Data <- read.csv("Plant_1_Generation_Data.csv")
Plant_2_Generation_Data <- read.csv("Plant_2_Generation_Data.csv")
Plant_1_Generation_Data <- Plant_1_Generation_Data[650:1651,]
Plant_2_Generation_Data <- Plant_2_Generation_Data[650:1651,]
```

Packages
```{r}
require('entropy')
require('nonlinearTseries')
require('crqa')
require('plot3D')
require('tseriesChaos')

```

estimate the delay parameter (The first local minimum provides the lag where the time-series is most independent of itself, and embedding at this point will provide the most unique information for  a  new  dimension  in  phase-space.) d1 = 7 d2 = 3
```{r}
mutual(Plant_1_Generation_Data$DC_POWER,lag.max=50)
mutual(Plant_2_Generation_Data$DC_POWER,lag.max=50)
```

estimate the embedding  dimension (m1 = 7, m2 = 8 m3=1)
```{r}
plot(false.nearest(Plant_1_Generation_Data$DC_POWER,m=10,d=7,t=0))
plot(false.nearest(Plant_2_Generation_Data$DC_POWER,m=10,d=2,t=0))
```


```{r}
Plant_1_Generation_Data <- Plant_1_Generation_Data[650:1651,]
Plant_2_Generation_Data <- Plant_2_Generation_Data[650:1651,]
```
Optimize parameters
```{r}
par =list(method ="crqa", metric ="euclidean", maxlag =20, radiusspan =100, radiussample =40, normalize =0, rescale =4, mindiagline =10, minvertline =10, tw =0, whiteline =FALSE, recpt =FALSE, side ="both", datatype ="continuous", fnnpercent  =20,  typeami ="mindip", nbins  =50, criterion ="firstBelow",threshold =1, maxEmb =20, numSamples =500, Rtol =10, Atol =2)
optParams <-optimizeParam(Plant_1_Generation_Data$DC_POWER, Plant_2_Generation_Data$DC_POWER, par,min.rec =2, max.rec =5)
```

CRQA analyse met d=7
```{r}
plant_result <- crqa (ts1 = Plant_1_Generation_Data$DC_POWER, ts2 = Plant_2_Generation_Data$DC_POWER,                  delay = 7, embed = 5, rescale =2, radius = 0.1019038, normalize = 2, 
                 mindiagline =2 , minvertline =2 , tw = 0 , whiteline = FALSE, recpt = FALSE, side =                  "both", method = "crqa", metric = "euclidean", datatype = "continuous")
```

CRQA Analyse met d=16
```{r}
plant_result2 <- crqa (ts1 = Plant_1_Generation_Data$DC_POWER, ts2 = Plant_2_Generation_Data$DC_POWER,delay = 16, embed = 5, rescale =2, radius = 0.1019038, normalize = 2, mindiagline =2 , minvertline =2 , tw = 0 , whiteline = FALSE, recpt = FALSE, side =                  "both", method = "crqa", metric = "euclidean", datatype = "continuous")
```


Resultaten en plot CRQA
```{r}
print(plant_result[1:9])
RP =as.matrix(plant_result$RP)
image(RP)

```



DCRP
```{r}
df <- data.frame(Plant_1_Generation_Data$DC_POWER,Plant_2_Generation_Data$DC_POWER)
#transform the data in a data.frame
normalizeddf <- as.data.frame(scale(df)) #normalize the variables keeping the data.frame format
res = drpfromts(normalizeddf$Plant_1_Generation_Data.DC_POWER, normalizeddf$Plant_2_Generation_Data.DC_POWER, windowsize = 20,
            radius = 0.001, delay = 1, embed = 1, rescale = 0,
             normalize = 0, mindiagline = 2, minvertline = 2,
             tw = 0, whiteline = FALSE, recpt = FALSE, 
             side = 'both', method = 'crqa', 
            metric = 'euclidean', datatype = 'continuous')
profile = res$profile 
plot(-100:100, profile, type ="l", lwd =2, xlab ="Lag", ylab ="%REC")
