---
title: "Module 2"
output:
  html_document: default
  pdf_document: default
---
load libraries and dataset

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(forecast)
library(fpp)
library(fpp2)

dataDIR <- "C:/Users/Gebruiker/Dropbox/Uvt/01 Master/Jaar 1/P3/Complex systems/Data/Input/"

globaltemperatures <- read.csv(paste0(dataDIR, "GlobalTemperatures.csv"))

#import dataset
#globaltemperatures <- read.csv("input/GlobalTemperatures.csv", stringsAsFactors = FALSE)
#globaltemperatures$observation <- 1:nrow(globaltemperatures) 
globaltemperatures$dt <- as.Date(globaltemperatures$dt,
                                   "%Y-%d-%m",
                                   tz = "Europe/London")
```

last 260 years in trends
```{r}
glob_land_1750_2015 <- window(globaltemperatures$LandAverageTemperature, start = 1, end = 3192)
temp_1750_1882 <- ts(glob_land_1750_2015, frequency = 12, start = c(1750,1))
glob_land_ma_1750_2015 <- ma(glob_land_1750_2015, order = 12)
Temp_c.1 <- ts(glob_land_ma_1750_2015, frequency = 12, start = c(1750,1))

ggplot(data = globaltemperatures, aes(x = dt, y = glob_land_1750_2015))+ geom_line() +
  theme_bw()+
  labs(title = "Temperature 1750-2015",
       x = "Year",
       y = "Temperature")

ggplot(data = globaltemperatures, aes(x = dt, y = glob_land_1750_2015))+ geom_line() +
  geom_line(aes(dt,glob_land_ma_1750_2015),color="red") + 
  theme_bw()+
  labs(title = "Temperature 1750-2015",
       x = "Year",
       y = "Temperature")

```

Regression line through moving average
```{r}
ggplot(data = globaltemperatures, aes(x = dt, y = glob_land_ma_1750_2015))+ geom_line()+
  theme_bw()+
  geom_smooth(method='lm', formula= y~x, color = "red") + 
  labs(title = "Temperature 1750-2015",
       x = "Year",
       y = "Temperature")
```

ggplot(data = globaltemperatures, geom_line(aes(x = Time, y = Landtemperatures))+
  geom_smooth(method='lm', formula= y~x, color = "red") + 
  labs(title = "Temperature 1750-2015",x = "Year", y = "Temperature")
  theme_bw()

1750-1882
```{r}
glob_land_1750_1882 <- window(globaltemperatures$LandAverageTemperature, start = 1, end = 1597)
glob_land_ma_1750_1882 <- ma(glob_land_1750_1882, order = 12)
Temp_c.2 <- ts(glob_land_ma_1750_1882, frequency = 12, start = c(1750,1))
globaltemperatures1 <- globaltemperatures[1:1597,]

ggplot(data = globaltemperatures1, aes(x = dt, y = glob_land_1750_1882))+
  geom_line() +
  geom_line(aes(x=dt,y=glob_land_ma_1750_1882),color="red") +
  theme_bw()+
  labs(title = "Temperature 1750-1882",
       x = "Year",
       y = "Temperature")
```

1882-1970
```{r}
glob_land_1882_1970 <- window(globaltemperatures$LandAverageTemperature, start = 1597, end = 2677)
glob_land_ma_1882_1970 <- ma(glob_land_1882_1970, order = 12)
Temp_c.3 <- ts(glob_land_ma_1882_1970, frequency = 12, start = c(1883,1))
globaltemperatures2 <- globaltemperatures[1597:2677,]

ggplot(data = globaltemperatures2, aes(x = dt, y = glob_land_1882_1970))+
  geom_line() +
  geom_line(aes(x=dt,y=glob_land_ma_1882_1970),color="red") +
  theme_bw()+
  labs(title = "Temperature 1882-1970",
       x = "Year",
       y = "Temperature")

```

```{r}
glob_land_1973_2015 <- window(globaltemperatures$LandAverageTemperature, start = 2677, end = 3192)
glob_land_ma_1973_2015 <- ma(glob_land_1973_2015, order = 12)
Temp_c.4 <- ts(glob_land_ma_1973_2015, frequency = 12, start = c(1973,1))
globaltemperatures3 <- globaltemperatures[2677:3192,]

ggplot(data = globaltemperatures3, aes(x = dt, y = glob_land_1973_2015))+
  geom_line() +
  geom_line(aes(x=dt,y=glob_land_ma_1973_2015),color="red") +
  theme_bw()+
  labs(title = "Temperature 1973-2015",
       x = "Year",
       y = "Temperature")
```


ggplot(data = globaltemperatures, geom_line(aes(x = Time, y = Landtemperature))+
  geom_line(aes(x = Time, y = Landtemperature_movingaverage), color = "red") +
  labs(title = "Temperature 1973-2015",x = "Year",y = "Temperature")
  theme_bw()

       
       
       
stl function
```{r}
#periodic behaviours
#1755-1770
glob_land_1755_1770 <- window(globaltemperatures$LandAverageTemperature, start = 61, end = 252)
temp_1755_1770 <- ts(glob_land_1755_1770, frequency = 12, start = c(1755,1))
glob_land_1755_1770 <- na.omit(glob_land_1755_1770)
z.1 <- ts(glob_land_1755_1770, frequency = 12, start = c(1755,1))
plot(stl(z.1, "periodic"),main = "STL plot 1755-1770")

plot(temp_1755_1770)
glob_land_ma_1755_1770 <- ma(temp_1755_1770, order = 12)
lines(glob_land_ma_1755_1770, col =2)
```

#1890-1905
```{r}
glob_land_1890_1905 <- window(globaltemperatures$LandAverageTemperature, start = 1681, end = 1860)
temp_1890_1905 <- ts(glob_land_1890_1905, frequency = 12, start = c(1890,1))
plot(temp_1890_1905)
glob_land_ma_1890_1905 <- ma(temp_1890_1905, order = 12)
lines(glob_land_ma_1890_1905, col =2)

glob_land_1890_1905 <- na.omit(glob_land_1890_1905)
z.2 <- ts(glob_land_1890_1905, frequency = 12, start = c(1890,1))
plot(stl(z.2, "periodic"), main="STL plot 1890-1905")
```

```{R}
#2000-2015 
glob_land_2000_2015 <- window(globaltemperatures$LandAverageTemperature, start = 3001, end = 3192)
temp_2000_2015 <- ts(glob_land_2000_2015, frequency = 12, start = c(2000,1))
plot(temp_2000_2015)
glob_land_ma_2000_2015 <- ma(temp_2000_2015, order = 12)
lines(glob_land_ma_2000_2015, col =2)

glob_land_2000_2015 <- na.omit(glob_land_2000_2015)
z.3 <- ts(glob_land_2000_2015, frequency = 12, start = c(2000,1))
plot(stl(z.3, "periodic"),main="STL plot 2000-2015")
```